---
title: "DVM Permutation Test - Work 2"
author: "Jeremy Williams and Spyridoula Chrysikopoulou-Soldatou"
date: "October 18, 2017"
output:
  pdf_document: default
  html_document: default
---

## **Prediction of Multiple Regression Model**

```{r}
wc <- c(16,20,25,27,32,48,48)
t <- c(75,83,85,85,92,97,99)
tmg <- c(1.85,1.25,1.5,1.75,1.15,1.75,1.6)

# Fit a linear model and run a summary of its results.
mod1<-lm(wc ~ t + tmg)
summary(mod1)

# Predicted values
fitted(mod1) 

# Model coefficients
coefficients(mod1)

# CIs for model parameters
confint(mod1, level=0.95)

# Plot model residuals.
plot(mod1, pch=16, which=1)

# Extract coefficients
mod1_coef <- round(coef(mod1), 3) 

# Display equation 
mtext(bquote(wc == .(mod1_coef[3])* tmg +.(mod1_coef[2])* t + .(mod1_coef[1])),
      adj=1, padj=-2) 
```

******

## **Significance of the Variables T and TMG**

We begin by testing whether a set of independent variables has   no   partial   effect   on   the   dependent   variable, "Y". 

Our model is: 

**wc = B0 + B1t + B2tmp + e**

**Null Hypothesis**: The initial assumption is that there is no relation, which is expressed as: 

**Ho: B1 = B2 = 0**.

**Alternative Hypothesis**: At least one of the independent variables IS useful in explaining/predicting Y, expressed as: 

**H1: At least one Bi is "not equal to" 0**.

### **Exact Permuatition Test **

```{r}
#install.packages("combinat")
suppressMessages(suppressWarnings(library("combinat")))
wc1 <- c(16,20,25,27,32,48,48)
t1 <- c(75,83,85,85,92,97,99)
tmg1 <- c(1.85,1.25,1.5,1.75,1.15,1.75,1.6)
mod2<-glm(wc1 ~ t1 + tmg1, family = gaussian)
a<-summary(mod2)
Tinttrue<-abs(a$coefficients[1,3])
Tttrue<-abs(a$coefficients[2,3])
Ttmgtrue<-abs(a$coefficients[3,3])

#number of rearrangements to be examined
n<-length(wc1)
nr<-fact(n) 
nr

Tint=numeric(nr);Tt=numeric(nr);Ttmg=numeric(nr)
newy<-permn(wc1)
for (i in 1:nr){
  mod2<-glm(newy[[i]] ~ t1 + tmg1,family = gaussian)
  a<-summary(mod2)
  Tint[i]<-abs(a$coefficients[1,3]) 
  Tt[i]<-abs(a$coefficients[2,3])
  Ttmg[i]<-abs(a$coefficients[3,3])}

par(mfrow=c(1,1))
hist(Tint)
abline(v=Tinttrue, lwd=4, col="blue")

#True t-value of intercept "B0"
Tinttrue

#P-Value of intercept "B0"
length(Tint[Tint>= Tinttrue])/nr

par(mfrow=c(1,1))
hist(Tt)
abline(v=Tttrue, lwd=4, col="blue")

#True t-Value of Time (T) "B1"
Tttrue

#P-value of Time (T) "B1"
length(Tt[Tt>= Tttrue])/nr

par(mfrow=c(1,1))
hist(Ttmg)
abline(v=Ttmgtrue, lwd=4, col="blue")

#True t-Value of Time moving the grass (TMG) "B2"
Ttmgtrue

#P-Value of Time moving the grass (TMG) "B2"
length(Ttmg[Ttmg>= Ttmgtrue])/nr
```

All (intercept, T and TMG) small p-values (p "less than or equal to" 0.05) indicates strong evidence against the null hypothesis, so we also reject the null hypothesis (Ho is rejected).

### **Multiple Regression through the Origin**

```{r}
 
#install.packages("combinat")
suppressMessages(suppressWarnings(library("combinat")))
#install.packages("ape")
suppressMessages(suppressWarnings(library("ape")))
wc1 <- c(16,20,25,27,32,48,48)
t1 <- c(75,83,85,85,92,97,99)
tmg1 <- c(1.85,1.25,1.5,1.75,1.15,1.75,1.6)
a5<-data.frame(wc1,t1,tmg1)

#number of rearrangements to be examined
nr<-fact(length(wc1))
nr 

#Permutation method using number of rearrangements
lmorigin(wc1 ~ t1 + tmg1, data =a5, nperm=nr) 
```

Both (parametric and permutational) small p-values (p "less than or equal to" 0.05) indicates strong evidence against the null hypothesis, so we also reject the null hypothesis (Ho is rejected).

### **LM and GLMs Comparison** 

```{r}

#install.packages("texreg")
suppressMessages(suppressWarnings(library("texreg")))
wc1 <- c(16,20,25,27,32,48,48)
t1 <- c(75,83,85,85,92,97,99)
tmg1 <- c(1.85,1.25,1.5,1.75,1.15,1.75,1.6)

# Estimate with OLS (Model 1):
reg1<-lm(wc1 ~ t1 + tmg1)

# Estimate with GLS (Model 2):
reg2<-glm(wc1 ~ t1 + tmg1, family = gaussian)

# Compare:
screenreg(l = list(reg1,reg2))
```

Both (LM and GLM) small p-values (p "less than or equal to" 0.05) indicates strong evidence against the null hypothesis, so we also reject the null hypothesis (Ho is rejected).

******

## **Conclusion**

After testing the significance of the variables T and TMG, we can conclude that at least one of the variables "T" and "TMG"" are significant. 



