---
title: "DVM Parametric Bootstrap - Homework 3"
author: "Jeremy Williams and Spyridoula Chrysikopoulou-Soldatou"
date: "October 25, 2017"
output:
  pdf_document: default
  html_document: default
---

## **Multiple capture-recapture problem**

### **Chao's estimator (No. of females)**

```{r}
install.packages("SPECIES")
library(SPECIES)

F_sights <- data.frame(sights = 1:7, # Number of sight cases, 
                       nbear = c(11, 13, 5, 1, 1, 0, 2)) # Frequencies

F_sights

chao1984(F_sights,conf=0.95)

```

This returns: 

**Nhat** - point estimate of the no. of Females

**SE** - standard error of the point estimate

**CI** - confidence interval using a log transformation explained in Chao 1987.

Note: We can also show this below by estimating the Lambda and then Parametric bootstrap (to compute CI).


### **Zero-truncated Poisson Distribution (Lambda estimate and CI)**


```{r}

#install.packages("SPECIES")
library(SPECIES)

F_sights <- data.frame(sights = 1:7, # Number of sight cases, 
                       nbear = c(11, 13, 5, 1, 1, 0, 2)) # Frequencies

PermF_T <- chao1984(F_sights,conf=0.95)$Nhat

#install.packages("fossil")
library(fossil)


Fu <- function(x) x/(1-exp(-x))-2.3
Ux <- uniroot(Fu,c(1,10))

Lamd <-Ux$root


# Parametric bootstrap (compute CI)
set.seed(11)
nsim <-1000
PermF <-numeric((length (nsim)))
     for (i in 1:nsim){ 
       n<- rpois(33,Lamd)
       while (any(n==0)== "TRUE"){
         n <-rpois(33,Lamd)
       }
       PermF[i] <-chao2(n)
     }

#Lambda
Lamd

quantile(PermF, probs = c(0.025, 0.975))

par(mfrow=c(1,1))
hist(PermF)
abline(v=PermF_T, lwd=4, col="blue")

```

It is estimated that lambda is **1.983573**. 

The blue line shows "Nhat" (**38**) - point estimate of the no. of Females

The confidence interval of the total number of females is (**34.19841**, **51.75000**) 
