---
title: 'DVM Reproducible Research'
author: "Jeremy Williams"
date: "October 6, 2017"
output:
  pdf_document: default
  html_document: default
---

## **CO2 Uptake in Grass Plants**

*This document provides a reproducible research of "Carbon Dioxide Uptake in Grass Plants" from a library dataset in "R" called "datasets".* 

*The research was complied into a initial data frame called "CO2"*

*CO2 has 84 rows and 5 columns of data from an experiment on the cold tolerance of the grass species "Echinochloa crus-galli".*

*CO2 was uptaken within six (6) plants from Quebec and six (6) plants from Mississippi.* 

*It was measured at several levels of ambient CO2 concentration. Half the plants of each type were chilled overnight before the experiment was conducted.*

******

## **Summary of CO2**

```{r}
#install.packages("datasets")
suppressMessages(suppressWarnings(library(datasets))) 
data(CO2)
summary(CO2)
```

## **Summary Explained**

- **Plant**: *an ordered factor with levels **Qn1 < Qn2 < Qn3 < ... < Mc1** giving a unique identifier for each plant.*

- **Type**: *a factor with levels Quebec and Mississippi giving the origin of the plant*

- **Treatment**: *a factor with levels nonchilled chilled*

- **Conc**: *a numeric vector of ambient carbon dioxide concentrations (mL/L)*.

- **Uptake**: *a numeric vector of carbon dioxide uptake rates (umol/m^2 sec)*.

******
## **Relationship between "Type" and "Treatment"**

### **Table of CO2-Type **
```{r}
table(CO2$Type)
```

### **Table of CO2-Treatment **
```{r}
table(CO2$Treatment)
```

### **Table of CO2-Type and Treatment ** 
```{r}
table(CO2$Type,CO2$Treatment)
```

******

## **Cross Table of CO2 - Type and Treatment**

### **Relationship of Type and Treatment using the Cross Tabulations**


```{r}
#install.packages("gmodels")
suppressMessages(suppressWarnings(library("gmodels")))
#Contingency tables 
joint.CO2<-CrossTable(CO2$Type,CO2$Treatment, prop.chisq = F)
```
- **Note**: *Based on PROC Frequency "PROC FREQ" in SAS or "CROSSTABS" in SPSS"*

### **Nested Contingency Tables** 
```{r}
joint.CO2
```
- **Note**: *Based on PROC Frequency "PROC FREQ" in SAS or "CROSSTABS" in SPSS"*

******
## **Barplot Relationship of "Type" and "Treatment" **

```{r}
par(mfrow=c(1,1))
joint.CO2_count<-joint.CO2$t
barplot(joint.CO2_count, beside = TRUE, col = rainbow(2), 
        ylab = "Type", xlab = "Treatment")
legend("center", c("Quebec", "Mississippi"), 
       pch = 15, col = rainbow(2))
```

******
## **"Conc" and "UpTake" using Histograms**

### **Histogram of CO2 Concentrations**
```{r}
par(mfrow=c(1,1))
hist(CO2$conc, xlab="CO2 Concentrations", 
     main="Histogram of CO2 Concentrations", col="red")
```

### **Histogram of CO2 Concentrations with it's Density Curve (DC)**
```{r}
par(mfrow=c(1,1))
x2 <- CO2$conc
hist(CO2$conc, xlab="CO2 Concentrations w/ DC", 
     main="Histogram of CO2 Concentrations w/ DC", 
     col="green", probability=TRUE)
lines(density(x2),col="blue", lwd=2)
```

### **Histogram of CO2 Uptake Rates **
```{r}
par(mfrow=c(1,1))
hist(CO2$uptake, xlab="CO2 Uptake Rates", 
     main="Histogram of CO2 Uptake Rates", col="red")
```

### **Histogram of CO2 Uptake Rates with it's Density Curve (DC)** 
```{r}
par(mfrow=c(1,1))
x3 <- CO2$uptake
hist(CO2$uptake, xlab="CO2 Uptake Rates w/ DC", 
     main="Histogram of CO2 Uptake Rates w/ DC", 
     col="green", probability=TRUE)
lines(density(x3),col="blue", lwd=2)
```

******
## **Linear Regression of "Conc" and "Uptake"**

### **Linear Regression Analysis of "Conc" and "Uptake"**
```{r}
fm <- lm(CO2$uptake ~ CO2$conc, data=CO2)
summary(fm)
```

### **Linear Regression Model of "Conc" and "Uptake"**
```{r}
par(mfrow=c(1,1))
plot(CO2$conc, CO2$uptake, 
     xlab="Ambient carbon dioxide concentrations (mL/L)", 
     ylab="Carbon dioxide uptake rates (umol/m^2 sec)", 
     main="Linear Regression Model of Conc and Uptake", col="red")
fm <- lm(CO2$uptake ~ CO2$conc, data=CO2)
abline(coef(fm), lty=4, col="blue")

#Extract coefficients of Conc and Uptake
fm_coef <- round(coef(fm), 3) 

#Display the regression equation of the line
mtext(bquote(y == .(fm_coef[2])*x + .(fm_coef[1])),
      adj=1, padj=0)
```
